---
title: "p8105_hw1_lz2657"
author: Lingyu Zhang
date: Sep 19, 2018
output: github_document
---

```{r tidyverse, include = FALSE}
library(tidyverse)
```

# Problem 1

First, I create the required data frame:

```{r problem1_df}
set.seed(1)

problem1_df = tibble(
  vec_unif = runif(n = 10, min = 0, max = 5),
  vec_logical = vec_unif > 2,
  vec_char = c("a","b","c","d","e","f","g","h","i","j"),
  vec_factor = factor(c("right","left","right","right","left","right","right","left","right","left"))
)

problem1_df
```

Now I take the mean of each variable in the dataframe:

```{r problem1_mean, eval = FALSE}
mean(problem1_df$vec_unif)
mean(problem1_df$vec_logical)
mean(problem1_df$vec_char)
mean(problem1_df$vec_factor)
```

The means of character and factor variables don't work because they are not numerical.

Then try to convert logical, character, and factor variables to numeric:

```{r problem1_numeric, eval = FALSE}
as.numeric(problem1_df$vec_logical)
as.numeric(problem1_df$vec_char)
as.numeric(problem1_df$vec_factor)
```

We can see that logical and factor variables can be converted while the charater ones can not.

Now try to convert the character variable to factor to numeric and the factor variable to character to numeric:

```{r problem1_doubleconvert}
as.numeric(as.factor(problem1_df$vec_char))

as.numeric(as.character(problem1_df$vec_factor))
```

It turns out that the former one works while the latter one doesn't because character variables can't be converted to numeric ones.

# Problem 2

First, I create a required data frame:

```{r problem2_df}
set.seed(2)

problem2_df = tibble(
  x = rnorm(n = 1000, mean = 0, sd = 1),
  y = rnorm(n = 1000, mean = 0, sd = 1),
  vec_logical = x + y > 0,
  vec_num = as.numeric(vec_logical),
  vec_factor = as.factor(vec_logical)
)
```

The size of the dataset is `r dim(problem2_df)`. 

The mean and median of x is `r mean(problem2_df$x)`, `r median(problem2_df$x)`, respectively.  

The proportion of cases for which the logical vector is TRUE is `r sum(problem2_df$vec_logical)/1000`.